---
title: "Storm Data Analysis for Peer-graded Assignment"
author: "Caitlin"
date: "2022-12-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background of this analysis

This is data analysis regarding how storms and other severe weather events can cause both public health and economic problems. The dataset used in this analysis is from the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database, covering data from 1950 to November 2011. 

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# load data and have an overview
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, "stormdata.csv.bz2")
df_storm <- read.csv("stormdata.csv.bz2")

str(df_storm)
head(df_storm)

```

## Data processing

You can also embed plots, for example:

```{r pressure, echo=TRUE}
sum(is.na(df_storm$EVTYPE))
sum(is.na(df_storm$STATE))

sum(is.na(df_storm$FATALITIES))
sum(is.na(df_storm$INJURIES))
sum(is.na(df_storm))

df_storm1 <- df_storm %>% select(c("EVTYPE","FATALITIES","INJURIES","PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP")) %>% rename_all(tolower)

str(df_storm1 )

df_health <- df_storm1 %>%  select(evtype, fatalities, injuries) %>%  group_by(evtype) %>% summarize(fatalities = sum(fatalities), injuries = sum(injuries)) %>% arrange(desc(fatalities),desc(injuries)) %>% slice(1:10) %>% gather(key = type, value = value, fatalities, injuries)

str(df_health)

unique(df_storm1$propdmgexp)
unique(df_storm1$cropdmgexp)

cost <- function(x) {
  if (x == "H")
    1E-4
  else if (x == "K")
    1E-3
  else if (x == "M")
    1
  else if (x == "B")
    1E3
  else
    1-6
}

df_economic <- df_storm1 %>% select(evtype, propdmg, propdmgexp, cropdmg,cropdmgexp) %>% 
  mutate(prop_dmg = propdmg*sapply(propdmgexp, FUN = cost), crop_dmg = cropdmg*sapply(cropdmgexp, FUN = cost),.keep = "unused") %>% group_by(evtype) %>% summarize(property = sum(prop_dmg), crop = sum(crop_dmg), .groups='drop') %>% arrange(desc(property), desc(crop))  %>% slice(1:10) %>% gather(key = type, value = value, property, crop)
  
str(df_economic)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
## Plot the results to answer below two questions

1.Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?
    
2.Across the United States, which types of events have the greatest economic consequences?


```{r}
ggplot(data = df_health, aes(reorder(evtype, -value), value, fill = type))+
  geom_bar(position = "dodge", stat = "identity") + 
  labs(x = "Event Type", y = "Count")+ theme_bw() + 
  theme(axis.text.x = element_text(angle = 20, vjust = 0.7)) + 
  ggtitle("Total Number of Fatalities and Injuries of top 10 storm event types") + 
  scale_fill_manual(values = c("purple", "blue"))


ggplot(data=df_economic, aes(reorder(evtype, -value), value, fill=type)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x="Event Type", y="Count (millions)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, vjust=0.5)) + 
  ggtitle("Total Cost of Property and Crop Damage by top 10 storm event types") +
  scale_fill_manual(values=c("darkgreen", "grey"))
       
```

# Conclusion

To sum up, tornado is the most impactful event to human health. More resources should be arranged to prevent tornado to guarantee the public health. Flood and hurricane/typhoon cause the most damage to economy espesially for property. The government should better system and infrastructure to protect property and crop.
